Constraining Black Hole Growth with High-z Quasar Observations

### 2022-08-25 ms.tex
version0825; before Euclid from me.

On the other hand, studies on X-ray selected faint quasars report number densities much higher than that
extrapolated from rest-UV observations \citep{2015A&A...578A..83G,2019ApJ...884...19G}.
In addition to the systematic difference in the quasar identification methods,
the X-ray observation should suffer less dust obscuration (see \S~\ref{sec:LF}),
thus detect more quasars in the faint end.
We also show the intrinsic QLF predicted by our best-fit parameters,
which is without imposing obscuration thus representing the upper bound of detectable quasar numbers.
Our intrinsic QLF is broadly consistent with the faint-end QLF from X-ray observation,
despite slightly lower than their data points at $\Muv\gtrsim -21$ mag.
% outshining effect
From another point of view, studies explain the decrepancy of the UV and X-ray QLF in the faint end
by pointing out that only AGNs outshining their host galaxies can be identified by the point source criteria
in the rest-UV selection \citep{2020MNRAS.495.2135N,2022arXiv220712282K}.
Especially in the faint end at $\Muv\gtrsim -24$ mag, the galaxy population can potentially contribute to this contamination significantly 
\citep{2022ApJS..259...20H}.
We do not take into account of this effect in the QLF correction,
but we notice this important caveat and provide the galaxy number density information 
in the higher-$z$ QLF predictions (see \S~\ref{sec:qlfglf}).

@@

\subsection{Quasar luminosity function}
\label{sec:qlfglf}
Next, we show our prediction of $z=6-10$ unobscured QLFs taken from the BH growth model in the left panel of Fig.~\ref{fig:QLF_rhoz}.
Similar to the BHMF, for all the redshift range, the shape of the QLF is well approximated with a double power-law function.
From higher to lower redshifts, the number density of luminous quasars increases and the bright-end of the distribution rises.
% %The model QLFs we propose give interesting prediction and will be further constrained by higher redshift quasar observations.
% \red{KI: add more explanations about the QLFs. How about the detection limits of Euclid, Roman, and others?
% Also, you could compare your results with Aaron Yung's results.}
In the right panle of Fig.~\ref{fig:QLF_rhoz}, we present the number density of luminous quasars with $\Muv <-26~{\rm mag}$ as a function of redshift.
The evolutionary trend can be fitted as $\propto 10^{k_{\rm L}z}$, where $k_{\rm L}\approx -0.97$.
A rapid decay of the quasar number density toward higher redshifts is found by previous studies \citep[e.g.,][]{2001AJ....122.2833F,2016ApJ...833..222J},
which show $k_{\rm L}= -0.47$ and $-0.72$ from the analysis of quasar samples at $z\simeq 3-5$ and $z\simeq 5-6$, respectively.
Recently, a steeper decline of luminous quasar number density toward higher redshifts is proposed by \citet{2019ApJ...884...30W},
suggesting $k_{\rm L}=-0.78$ at $6\lesssim z \lesssim 7$.
%a steeper decline ifts is suggested in recent studies of $z\gtrsim 6$ quasars,
%e.g. \citet{2019ApJ...884...30W} measure the spatial density of $z\sim 6.7$ quasars,
%and fit the parameter $k_{\rm L}=-0.78$ by connecting the $z\sim 6$ result from \citet{2016ApJ...833..222J}.
The rapid decline toward higher redshifts is characterized by the treatment of episodic natures of BH accretion with a finite value of
$\tau ~(\simeq 20~{\rm Myr})$, in order to match the statistical properties of high-$z$ quasars.
We also note that the tendency would be explained by the rarity of massive quasar-host halos at higher redshifts,
non-constant radiative efficiencies that covert the BH growth rate into the luminosity, or some combination of these effects
\citep[e.g.,][]{2010ApJ...718..231S}.
%Studies with the assumption of the close correlation of DM halo mass with central BH mass, and a constant Eddington ratio 
%interprete this rapid decline by the rarity of massive halos toward higher redshifts \citep[e.g.,][]{2010ApJ...718..231S}, 
%while another entangling factor is the radiative efficiency in converting BH growth rate to luminosity. 
%In our work, we argue that the episodical BH accretion with bursts and quiescent phases is 
%important in matching the $z\sim 6$ quasar observations, 
%and this calls consideration of the more realistic BH-galaxy coevolution in understanding the observable AGN properties.

%On the other hand, the statistics of BH masses together with the ERDF in their growth history can rebuild quasar luminosity functions across cosmic time,
%which is interesting to be probed with future observations.
In Fig.~\ref{fig:LFs}, we divide the cosmological QLF into four different epochs; (a) $z\simeq 7$, (b) $z\simeq 8$, (c) $z\simeq 9$, and (d) $z\simeq 10-13$.
The black solid and dashed curves present the QLF for the unobscured population (i.e., type-I AGNs, same as those in Fig.~\ref{fig:QLF_rhoz})
and the whole population (i.e., type-I and type-II AGNs).
The AGN obscured fraction based on X-ray samples \citep{2014ApJ...786..104U} is applied to faint objects, 
but the correction factor is highly uncertain at the faint end of $\Muv\gtrsim -20$ mag.
% sampled in \citet{2014ApJ...786..104U},
%we emphasize that the observed (Type I) QLF after the current obscuration prescription can be highly uncertain in the faint end,
%ßwhich is to be improved by the future equipments.
% More thorough studies on the high-$z$ quasar obscuration by upcoming equipments will improve the understanding on this aspect.
%We also produce the intrinsic (Type I and Type II) luminosity function without obscuration effects.
%The QLF emgerging from high redshifts quickly evolves toward brighter luminosity 
%with the number of quasars increases as BHs grow.
We also overlay the expected depths and survey volumes of upcoming near-infrared surveys by
Euclid \citep{2011arXiv1110.3193L} and RST \citep{2019arXiv190205569A}.
Only their Wide survey layers are considered here.
Table.~\ref{tab:N_detect} summarizes the expected number of detected quasars in each survey. 
Euclid will identify quasars up to $z\sim 8$ and discover $O(100-1000)$ quasars at $z\simeq 7-8$.
The RST will be able to explore the very faint ($\Muv>-22$ mag) but abundant quasar population even at $z>10$, in principle,
after subtracting galaxy contaminants.

>>>>

On the other hand, studies on X-ray selected faint quasars report number densities much higher than that
extrapolated from rest-UV observations \citep{2015A&A...578A..83G,2019ApJ...884...19G}.
In addition to the systematic difference in the quasar identification methods,
the X-ray observation should suffer less dust obscuration (see \S~\ref{sec:LF}),
thus detect more quasars in the faint end.
We also show the intrinsic QLF predicted by our best-fit parameters,
which is without imposing obscuration thus representing the upper bound of detectable quasar numbers.
Our intrinsic QLF is broadly consistent with the faint-end QLF from X-ray observation,
despite slightly lower than their data points at $\Muv\gtrsim -21$ mag.
% outshining effect
From another point of view, studies explain the decrepancy of the UV and X-ray QLF in the faint end
by pointing out that only AGNs outshining their host galaxies can be identified by the point source criteria
in the rest-UV selection \citep{2020MNRAS.495.2135N,2021MNRAS.502.2757O,2022arXiv220712282K}.
Especially in the faint end at $\Muv\gtrsim -24$ mag, the galaxy population can potentially contribute to this contamination significantly 
\citep{2022ApJS..259...20H}.
We do not take into account of this effect in the QLF correction,
but we notice this important caveat and provide the galaxy number density information 
in the higher-$z$ QLF predictions (see \S~\ref{sec:qlfglf}).

@@

\subsection{Quasar luminosity function}
\label{sec:qlfglf}
Next, we show our prediction of $z=6-10$ unobscured QLFs taken from the BH growth model in the left panel of Fig.~\ref{fig:QLF_rhoz}.
Similar to the BHMF, for all the redshift range, the shape of the QLF is well approximated with a double power-law function.
From higher to lower redshifts, the number density of luminous quasars increases and the bright-end of the distribution rises.
% %The model QLFs we propose give interesting prediction and will be further constrained by higher redshift quasar observations.
% \red{KI: add more explanations about the QLFs. How about the detection limits of Euclid, Roman, and others?
% Also, you could compare your results with Aaron Yung's results.}
In the right panle of Fig.~\ref{fig:QLF_rhoz}, we present the number density of luminous quasars with $\Muv <-26~{\rm mag}$ as a function of redshift.
The evolutionary trend can be fitted as $\propto 10^{k_{\rm L}z}$, where $k_{\rm L}\approx -0.97$.
Observationally, a rapid decay of the quasar number density toward higher redshifts is found by previous studies
\citep[e.g.,][]{2001AJ....122.2833F,2013ApJ...768..105M,2016ApJ...833..222J,2019ApJ...884...30W},
which show $k_{\rm L}= -0.47$ and $-0.72$ from the analysis of quasar samples at $z\simeq 3-5$ and $z\simeq 5-6$, respectively.
% Recently, a steeper decline of luminous quasar number density toward higher redshifts is proposed by \citet{2019ApJ...884...30W},
% suggesting $k_{\rm L}=-0.78$ at $6\lesssim z \lesssim 7$.
The rapid decline toward higher redshifts is characterized by the treatment of episodic natures of BH accretion with a finite value of
$\tau ~(\simeq 20~{\rm Myr})$, in order to match the statistical properties of high-$z$ quasars.
We also note that the tendency would be explained by the rarity of massive quasar-host halos at higher redshifts,
non-constant radiative efficiencies that covert the BH growth rate into the luminosity, or some combination of these effects
\citep[e.g.,][]{2010ApJ...718..231S}.
%Studies with the assumption of the close correlation of DM halo mass with central BH mass, and a constant Eddington ratio 
%interprete this rapid decline by the rarity of massive halos toward higher redshifts \citep[e.g.,][]{2010ApJ...718..231S}, 
%while another entangling factor is the radiative efficiency in converting BH growth rate to luminosity. 
%In our work, we argue that the episodical BH accretion with bursts and quiescent phases is 
%important in matching the $z\sim 6$ quasar observations, 
%and this calls consideration of the more realistic BH-galaxy coevolution in understanding the observable AGN properties.

%On the other hand, the statistics of BH masses together with the ERDF in their growth history can rebuild quasar luminosity functions across cosmic time,
%which is interesting to be probed with future observations.
In Fig.~\ref{fig:LFs}, we divide the cosmological QLF into four different epochs; (a) $z\simeq 7$, (b) $z\simeq 8$, (c) $z\simeq 9$, and (d) $z\simeq 10-13$.
The black solid and dashed curves present the QLF for the unobscured population (i.e., type-I AGNs, same as those in Fig.~\ref{fig:QLF_rhoz})
and the whole population (i.e., type-I and type-II AGNs).
The AGN obscured fraction based on X-ray samples \citep{2014ApJ...786..104U} is applied to faint objects, 
but the correction factor is highly uncertain at the faint end of $\Muv\gtrsim -20$ mag.
% sampled in \citet{2014ApJ...786..104U},
%we emphasize that the observed (Type I) QLF after the current obscuration prescription can be highly uncertain in the faint end,
%ßwhich is to be improved by the future equipments.
% More thorough studies on the high-$z$ quasar obscuration by upcoming equipments will improve the understanding on this aspect.
%We also produce the intrinsic (Type I and Type II) luminosity function without obscuration effects.
%The QLF emgerging from high redshifts quickly evolves toward brighter luminosity 
%with the number of quasars increases as BHs grow.
We also overlay the expected depths and survey volumes of upcoming near-infrared surveys by
Euclid \citep{2011arXiv1110.3193L} and RST \citep{2019arXiv190205569A}.
Only their Wide survey layers are considered here.
Table.~\ref{tab:N_detect} summarizes the coverage and limiting magnitudes in each survey,
and list the expected number of quasar detection.
Euclid will identify quasars up to $z\sim 8$ and discover $O(100-1000)$ quasars at $z\simeq 7-8$.
The RST will be able to explore the very faint ($\Muv>-22$ mag) but abundant quasar population even at $z>10$, in principle,
after subtracting galaxy contaminants.
